Precise Spiking Motifs in Neurobiological and Neuromorphic~Data

### 2022-12-28 brainsci-2069565_accepted.tex
moving figures 6 <-> 7 to respect apparition order
fixing caption

\begin{figure}[H]%[t!][]
%  \centering
  \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
  \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
  \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
  \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
    \caption{
      Detecting event-based motifs using spiking neurons with heterogeneous delays.
    { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups'' (PGs). 
    { (\textbf{b})}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay. 
    { (\textbf{c})}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}. % such that the addresses are the synapses of these neurons (which we therefore call here synapses).
    %Note that each afferent may be connected with multiple weights at different delays. The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses. %These PG motifs were efficiently learned by a logistic regression.
    %{\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
    { (\textbf{d})}~Knowing the %MDPI: Please change the (a) into (d) in picture. Fixed !
 results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic example, all spikes are now associated with a PG. 
     }
  \label{fig:THC} % TODO: citer biol cyb fig:THC
\end{figure}

@@

% 
%----------------------------%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Learning Synaptic Delays}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Training delays ~\citep{state_training_2019}
First, the original model by Izhikevich uses a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection. For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity--connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}. Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}. They were also extended to recurrent neuronal networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}. In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity. More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity, and these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.

>>>>

@@

\begin{figure}[H]%[t!][]
    \includegraphics[width=0.490\linewidth]{figures/THC_1a_k.pdf}
    \includegraphics[width=0.490\linewidth]{figures/THC_1c.pdf}
    \includegraphics[width=0.490\linewidth]{figures/THC_1b.pdf}
    \includegraphics[width=0.490\linewidth]{figures/THC_1a.pdf}
      \caption{
        Detecting event-based motifs using spiking neurons with heterogeneous delays.
      { (\textbf{a})}~Given a generic raster plot defined by a set of spikes occurring on specific neuron addresses and at specific times, one may consider that this information consists of the repeated occurrence of groups of precise motifs of spikes that define ``polychronous groups''. 
      { (\textbf{b})}~The generative model is defined by this set of motifs (here $4$ PGs) each defined by  synapses defined by weights (increasing with the radius of the black dots) with heterogeneous delay. 
      { (\textbf{c})}~Generalizing the core polychrony detection model (see Figure~\ref{fig:izhikevich}), one can define a layer of neurons that detect the identity and timing of these PGs~\citep{grimaldi_learning_2022}. % such that the addresses are the synapses of these neurons (which we therefore call here synapses).
      %Note that each afferent may be connected with multiple weights at different delays. The propagation of the afferent information through these delay may generate at each time step a synchronous motif on a subset of synapses. %These PG motifs were efficiently learned by a logistic regression.
      %{\bf (c)}~The output of the model provides with the predicted probability of occurrence of each PG motif at any time, which may be used to generate a spike as a Bernoulli trial, providing in this particular case with an exact identification of PGs occurrences.
      { (\textbf{d})}~Knowing the %MDPI: Please change the (a) into (d) in picture. Fixed !
   results of this detection, one may for illustration purposes highlight them by different colors in the raster plots, showing that in this synthetic example, all spikes are now associated with a PG. 
    \label{fig:THC} % TODO: citer biol cyb fig:THC
  \end{figure}
\subsection{Learning Synaptic Delays}
% Training delays ~\citep{state_training_2019}
First, the original model by Izhikevich uses a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection. For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity--connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}. Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}. They were also extended to recurrent neuronal networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}. In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity. More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity, and these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.
